### 2018-09-23 concord/data/table2.tex
Converted concord to package and updated scripts
A major rearrangement to allow concord to be structured as a package, and
installed using pip. This modification means that the $CONCORD_PATH
environemnt variable is no longer required; the code can find the data
files required via other methods

The anisotropy function is now part of the diskmodel subpackage.

Also updated the tutorial notebook to make sure the code still works after
all the changes. The data files required are now part of the repo, in the
example_data subdirectory

\begin{table*}[h]
\caption{Properties of thermonuclear bursts observed from target sources by
the {\it Rossi X-ray Timing Explorer}}\label{tab:bursts}
\begin{tabular}{lccccccccc}
Epoch & $N_{\rm burst}$ ($N_{\rm av}$) & burst IDs$^a$ & $\Delta t$ (hr) & $F_{\rm per}^b$ ($10^{-9}\ \epcs$) & $c_{\rm bol}$  & $\dot{m}$ ($\dot{m}_{\rm Edd}$) & $E_b$ ($10^{-6}\ \epc$) & $F_{\rm pk}$ ($10^{-9}\ \epcs$) & $\alpha$ \\
1998 Jun & 6(1) & 7 & $5.14\pm0.07$ & $1.167\pm0.006$ & $1.806\pm0.009$ & 0.0513 & $1.102\pm0.011$ & $30.9\pm1.0$ & $34.2\pm0.5$ \\ 
2000 Sep & 11(7) & 14--20 & $4.177\pm0.010$ & $1.593\pm0.017$ & $1.787\pm0.003$ & 0.0692 & $1.126\pm0.016$ & $29.1\pm0.5$ & $38.6\pm0.3$ \\ 
2007 Mar & 10(7) & 58--65 & $3.530\pm0.004$ & $1.87\pm0.02$ & $1.751\pm0.003$ & 0.0796 & $1.18\pm0.04$ & $28.4\pm0.4$ & $35.3\pm1.0$ \\ 
2002 Oct & 1(1) & 2 & $16.55\pm0.06^c$ & 2.541--2.298 & $2.085\pm 0.019$ & 0.0472$^d$ & $2.649\pm0.018$ & $229\pm4$ & $114.4\pm1.9$ \\ 
         & 1(1) & 3 & 21.10 & 2.298--1.946 & $2.13\pm0.04$ & 0.0432$^d$ & $2.990\pm0.017$ & $232\pm4$ & $118.2\pm1.9$ \\ 
         & 1(1) & 4 & 29.82 & 1.946--1.826 & $2.157\pm0.002$ & 0.0384$^d$ & $3.46\pm0.02$ & $232\pm4$ & $128.2\pm2.1$ \\ 
1997 May 4 & 3(1) & 1 & $2.681\pm0.007$ & $3.72\pm0.18$ & $1.45\pm0.09$ & 0.144 & $0.381\pm0.003$ & $61\pm2$ & $138.5\pm1.4$ \\
2009 Jun 12 & 2(2) & & $1.892^e$ & $5.70\pm0.04$ & 1.4981 & 0.226 & $0.371\pm0.010$ & $56.6\pm1.4$ & $160.1\pm1.8$\\ 
2001 Feb & 1(1) & & $\leq 1.53\times10^{4}$~$^f$ & $4.73\pm0.03$ & 1.5346 & 0.167 & $110\pm9$~$^g$ & $21.9\pm0.6$~$^g$ & \\
\hline
\end{tabular}
\medskip\\
$^a$ burst index number in the catalog of \cite{bcatalog}.\\
$^b$ persistent flux in the energy range 3--25 keV \\
$^c$ inferred from ignition model comparisons \\
$^d$ calculated based on linear interpolation between observations falling within the burst interval \\
$^e$ only two bursts detected, so may not have been part of a regular train \\
$^f$ recurrence time upper limit of 1.75~yr based on the next earliest event, detected by the \xte/ASM \cite[]{kuul04} \\
$^g$ values taken from the ``optimal'' fits of \cite{keek14a}

\end{table*}